---
title: "Missing Data Analysis: Full Information Maximum Likelihood"
author: "Butovens Médé"
date: "2/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
### Import Libraries
# install.packages("tidyverse", "skimr", "lavaan", "semTools", "kableExtra")
library(tidyverse)
library(skimr)
library(lavaan)
library(semTools)
library(kableExtra)
```


```{r}
### Load data
steroid_data <- read_csv("steroid.csv", col_names = T, na = "999")

### structure of data
steroid_data %>% skim()

### Subset data
steroid_data_small <- steroid_data %>% 
  select(mayuse,
         sterknow,
         testknow,
         suppknow,
         attuse,
         streneff,
         teaminfo,
         coachtol,
         esteem,
         impuls)

### Look at smaller data set
skim(steroid_data_small)

### Look at missing data pattern
VIM::aggr(steroid_data_small)
```


# 1: Estimate regression model using FIML in R
```{r}
### Build regression model for missing data (without auxiliary variable)
### Model Specification
model <- '
### Linear regression
mayuse ~ sterknow + testknow + suppknow

### Estimate variances of predictors with missing values to include all cases in analysis, and avoid listwise deletions
sterknow ~~ sterknow
testknow ~~ testknow
suppknow ~~ suppknow

### Estimate covariances of predictors with missing values to include all cases in analysis, and avoid listwise deletion
sterknow ~~ testknow
sterknow ~~ suppknow
testknow ~~ suppknow
'
### Model Estimation
modSEM <- sem(model = model,
              data = steroid_data_small,
              meanstructure = T,
              missing = 'fiml')

### Print/Extract results
summary(modSEM, rsquare = T)

### Standardized results (Only reported if specifically needed or asked)
# standardizedSolution(modSEM)
```


# 2: Estimate regression model using saturated correlates approach
```{r}
### Build regression model for missing data (with auxiliary variable)
### Model Specification
sat_model <- '
### Linear regression
mayuse ~ sterknow + testknow + suppknow

### Estimate variances of predictors with missing values to include all cases in analysis, and avoid listwise deletions
sterknow ~~ sterknow
testknow ~~ testknow
suppknow ~~ suppknow

### Estimate covariances of predictors with missing values to include all cases in analysis, and avoid listwise deletion
sterknow ~~ testknow
sterknow ~~ suppknow
testknow ~~ suppknow

### Correlate Auxiliary Variables with manifest (i.e. not latent) predictor variables
sterknow ~~ attuse + streneff + teaminfo + coachtol + esteem + impuls
testknow ~~ attuse + streneff + teaminfo + coachtol + esteem + impuls 
suppknow ~~ attuse + streneff + teaminfo + coachtol + esteem + impuls

### Correlate Auxiliary Variables with other Auxiliary Variables
attuse ~~ streneff
attuse ~~ teaminfo
attuse ~~ coachtol
attuse ~~ esteem
attuse ~~ impuls
streneff ~~ teaminfo
streneff ~~ coachtol
streneff ~~ esteem
streneff ~~ impuls
teaminfo ~~ coachtol
teaminfo ~~ esteem
teaminfo ~~ impuls
coachtol ~~ esteem
coachtol ~~ impuls
esteem ~~ impuls

### Correlate Auxiliary Variables with residual of the outcome
mayuse ~~ attuse + streneff + teaminfo + coachtol + esteem + impuls
'
### Model Estimation
sat_modSEM <- sem(model = sat_model,
              data = steroid_data_small,
              meanstructure = T,
              missing = 'fiml')

### Print/Extract results (unstandardized)
summary(sat_modSEM, rsquare = T)

### Standardized results (Only reported if specifically needed or asked)
# standardizedSolution(sat_modSEM)

##### Alternative method to include auxiliary variables automatically (using the sem.auxiliary fct from semTools)
aux_vars <- c("attuse", "streneff", "teaminfo", "coachtol", "esteem", "impuls")
sat_modSEM_auto <- sem.auxiliary(model = model,
              data = steroid_data_small,
              aux = aux_vars)
### Summary
summary(sat_modSEM_auto, rsquare = T)

### Standardized results (Only reported if specifically needed or asked)
# standardizedSolution(sat_modSEM_auto)
```


# 3: Table that contains regression coefficients and standard errors for model without and with AV:
```{r}
### Create data frame with needed values 
df <- data.frame(Predictors = c("sterknow", "testknow", "suppknow"),
           coef.R = c(-0.048, 0.058, -0.133),
           se.R = c(0.007, 0.052, 0.029),
           p.value = c(0.000, 0.264, 0.000),
           coef.MPlus = c(-0.048, 0.058, -0.133),
           se.MPlus = c(0.007, 0.052, 0.029),
           p.value = c(0.000, 0.0264, 0.000),
           coef.AV.R = c(-0.048, 0.062, -0.134),
           se.AV.R = c(0.007, 0.052, 0.029),
           p.value = c(0.000, 0.231, 0.000),
           coef.AV.MPlus = c(-0.048, 0.062, -0.134),
           se.AV.Mplus = c(0.007, 0.052, 0.029),
           p.value = c(0.000, 0.231, 0.004)
           )

### Create table
df %>% 
  kbl(caption = "Regression coefficients and standard errors for models without and with AVs using R and Mplus") %>% 
  kable_classic(full_width = F, html_font = "Cambria")
```

* Overall this model account for very little of the variance ($R^2 = 0.082$) seen in intention to use steroid (about 8%). The model without the auxiliary variables using FIML finds estimates to be $B_1 = -0.048$, $B_2= 0.058$, and $B_3= -0.133$ for steroid knowledge, testosterone effect knowledge and supplement knowledge respectively. Among these predictors, only steroid knowledge and supplement knowledge were significant ($p < 0.05$). More precisely the model found that for one unit of change in steroid knowledge, the intention to use **decreased** by 0.048 (everything else remaining constant). Similarly it was found that for one unit of change in supplement knowledge, the intention to use **decreased** by 0.133 (everything else remaining constant). These results were found using the software R.

* Identical results were found using Mplus. 

* When we compare the multiple regression model using FIML without the auxiliary variables to the one using the auxiliary variables, we notice slight changes in the estimation of the Beta coefficients (both in R and Mplus). While the estimation of Betas changed, no predictors switched from being significant to non-significant (and vice versa).

* Using the FIML method on the regression model with auxiliary variables, we also found that the only significant predictors of 'intention to use' were steroid knowledge and supplement knowledge. The estimate of the coefficient for steroid knowledge, $B_1$ was the same as the model without the auxiliary variables, and the estimate of the coefficient for the supplement knowledge, $B_3$ very slightly decreased to $-0.134$. Standard errors remained the same for all coefficients. The coefficient $B_2$ changed but remained non-significant.

* Identical results were found (compared to R) when including auxiliary variables in the MPlus model. 
